---
title: "CKME136 XJ0 - Step 2: Analyze data (non-textual) "
output: html_notebook
---

This is the second of four R Notebook files:

* CKME136 XJ0 - Step 1: Collect and pre-process data  
* **CKME136 XJ0 - Step 2: Analyze data (non-textual)**    
* CKME136 XJ0 - Step 3: Analyze data (textual)  
* CKME136 XJ0 - Step 4: Step 4: Machine learning for sentiment analysis  


##Pre-Processing

####Load the required packages

```{r load_packages}
#Load packages
library(rtweet) #Analyzing tweets 
library(tidyverse) #Wrangling data
library(tidytext) #Text processing
library(PerformanceAnalytics)
```

####Load filtered data sets
```{r load_data}
#Load filtered get_timelines data set
canpoli.tmls.filtered=read_csv("canpoli.tmls.filtered.csv") 

#Load filtered search_tweets data set 
canpoli.search.filtered=read_csv("canpoli.srch.filtered.csv") 
```

##Preliminary analysis 
```{r pre_analysis}
#View the summary of the dataframe
summary(canpoli.tmls.filtered)
#View the structure of the dataframe
str(canpoli.tmls.filtered)
```

##Analysis of variables (timelines)

####Tweets per CPPL (Canadian political party leader) and mentioned

```{r number_tweets}

#Overall CPPL tweets over duration of project
canpoli.tmls.filtered %>%
    ts_plot("days", trim = 1L) +
  ggplot2::geom_line(size=1) +
  ggplot2::theme_classic() +
  scale_color_manual("blue")+
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "bottom",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Tweets by CPPLs (original tweets & retweets)",
    subtitle = "Twitter status (tweet) counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Overall tweets mentioning CPPL over duration of project
options(scipen=10000)
canpoli.search.filtered %>%
  ts_plot("days", trim = 1L) +
  ggplot2::geom_line(size=1) +
  ggplot2::theme_classic() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "bottom",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Tweets mentioning CPPLs (original tweets & retweets)",
    subtitle = "Twitter status (tweet) counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Number of tweets per party leader
canpoli.tmls.filtered %>%
  ggplot(., aes(x = name,fill =is_retweet)) +
  geom_bar()+
  theme()+
    ggplot2::labs(
    x = NULL, y = NULL,
    title = "Total tweets by CPPLs (original tweets and retweets)",
    subtitle = "Tweets aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Number of tweets per party leader
canpoli.tmls.filtered %>%
  filter(is_retweet==FALSE) %>% 
  ggplot(., aes(x = name,fill =is_retweet)) +
  geom_bar()+
  theme()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Total tweets by CPPLs (original tweets)",
    subtitle = "Tweets aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )





```

####Tweets timeseries analysis

```{r timeseries}
#Number of tweets by party leader (comparison over time-original tweets and retweets)
canpoli.tmls.filtered %>%
  dplyr::group_by(name) %>%
  ts_plot("days", trim = 1L) +
  ggplot2::geom_line() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "bottom",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Tweets by CPPLs (original tweets & retweets)",
    subtitle = "Twitter status (tweet) counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Number of tweets by party leader (comparison over time original tweets only)
canpoli.tmls.filtered %>%
  dplyr::group_by(name) %>%
  dplyr::filter(is_retweet ==FALSE) %>% 
  ts_plot("days", trim = 1L) +
  ggplot2::geom_line() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "bottom",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Tweets by CPPLs (original tweets only)",
    subtitle = "Twitter status (tweet) counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Number of tweets by party leader (comparison over time - retweets only)
canpoli.tmls.filtered %>%
  dplyr::group_by(name) %>%
  dplyr::filter(is_retweet ==TRUE) %>% 
  ts_plot("days", trim = 1L) +
  ggplot2::geom_line() +
  ggplot2::theme_classic() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "bottom",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Tweets by CPPLs (retweets only)",
    subtitle = "Twitter status (tweet) counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

```

####Tweets source analysis

```{r source}
#Source of tweets by party leader
canpoli.tmls.filtered %>%
  ggplot(., aes(x = name, fill =source)) +
  geom_bar()+
  theme(legend.title = element_blank())+
    ggplot2::labs(
    x = NULL, y = NULL,
    title = "Sources of tweets by CPPLs  (original tweets & retweets)",
    subtitle = "Tweet counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )+
  coord_flip()

##Source of tweets by party leader (with retweets)
canpoli.tmls.filtered %>%
  ggplot(., aes(x = name, fill =source)) +
  geom_bar()+
  ggtitle("piped ggplot2")+
  theme(legend.title = element_blank())+
    ggplot2::labs(
    x = NULL, y = NULL,
    title = "Sources of tweets by CPPLs (original tweets vs. retweets)",
    subtitle = "Tweet) counts aggregated from February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )+ facet_wrap(~ is_retweet)+
  coord_flip()

```

####CPPLs followers analysis

```{r followers}
#Number of followers by CCPL (comparison)
options(scipen=10000)
canpoli.tmls.filtered %>%
  group_by(name) %>%
  summarize(avgfollowers = median(followers_count)) %>% 
  arrange(desc(avgfollowers)) %>% 
  ggplot(., aes(x = name, y=avgfollowers, fill=avgfollowers)) + 
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Number of followers of CPPLs",
    subtitle = "Twitter follower counts: February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )+
  coord_flip()

#All leaders excluding Justin Trudeau since Trudeau seems to be an outlier
canpoli.tmls.filtered %>%
  filter(name != "Justin Trudeau") %>% 
  group_by(name) %>%
  summarize(avgfollowers = median(followers_count)) %>% 
  arrange(desc(avgfollowers)) %>% 
  ggplot(., aes(x = name, y=avgfollowers, fill =avgfollowers)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Number of followers of CPPLs (excl. Trudeau)",
    subtitle = "Twitter follower counts: February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )+
  coord_flip()
```


####Tweets favourited analysis

```{r favourites}
#Number of tweets favourited by party leader (comparison)
#Count of tweets favourited
ggplot(canpoli.tmls.filtered, aes(x = name, y=favorite_count, fill=favorite_count)) +
  geom_bar(stat='identity') +
  theme(legend.title = element_blank())+
  #theme_classic()+
  #facet_wrap(~ source)+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Number of CPPLs tweets favourited",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )+
  coord_flip()

#Average of tweets favourited
canpoli.tmls.filtered %>% group_by(name) %>%     
  summarise(Number_of_Tweets = n_distinct(status_id), Number_of_Likes = sum(favorite_count)) %>% 
  mutate(Avg_Likes=Number_of_Likes/Number_of_Tweets)%>%
  ggplot(., aes(x = name, y=Avg_Likes, fill =Avg_Likes)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Average number favourites per CPPL tweet",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )+
  coord_flip()
```

####Tweet hashtag analysis (CPPL)

```{r cppl_hashtag}
#Hashtags by party leader (comparison)
#Create a seperate chart for each party leader to examine main themes

#Hastags by all CPPL (top 10)
canpoli.tmls.filtered %>%
  select(name, hashtags, is_retweet) %>% 
  filter(is_retweet ==FALSE) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(name,word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hashtags used by all CPPL",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )
  

#Hastags by Trudeau (top 10)
canpoli.tmls.filtered %>%
  filter(name == "Justin Trudeau") %>% 
  select(name, hashtags, is_retweet) %>% 
  filter(is_retweet ==FALSE) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(name,word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Hashtags used by Trudeau",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Hastags by Singh (top 10)
canpoli.tmls.filtered %>%
   filter(name == "Jagmeet Singh") %>% 
   select(name, hashtags, is_retweet) %>% 
  filter(is_retweet ==FALSE) %>%  
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(name,word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Hashtags used by Singh",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Hastags by Bernier (top 10)
canpoli.tmls.filtered %>%
  filter(name == "Maxime Bernier") %>% 
  select(name,hashtags, is_retweet) %>% 
  filter(is_retweet ==FALSE) %>%  
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(name,word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Hashtags used by Bernier",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Hastags by May (top 10)
canpoli.tmls.filtered %>%
  filter(name == "Elizabeth May") %>% 
  select(name, hashtags, is_retweet) %>% 
  filter(is_retweet ==FALSE) %>%  
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(name,word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Hashtags used by May",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Hastags by Scheer (top 10)
canpoli.tmls.filtered %>%
  filter(name == "Andrew Scheer") %>% 
  select(name, hashtags, is_retweet) %>% 
  filter(is_retweet ==FALSE) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(name,word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Hashtags used by Scheer",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

```

##Analysis of variables (search)
####Tweet hashtag analysis (Public)

```{r hashtag_buzz}
#Top 10 hastags directed at all party leaders
canpoli.search.complete %>%
  select(hashtags) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hashtags directed at CPPLs",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Top 10 hastags in tweets mentioning Trudeau
canpoli.search.complete %>%
   filter(grepl("@JustinTrudeau",text)) %>% 
  #filter(name == "Justin Trudeau") %>% 
  select(hashtags) %>% 
  #filter(is_retweet ==FALSE) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hastags in tweets mentioning Trudeau",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Top 10 hastags in tweets mentioning Bernier
canpoli.search.complete %>%
   filter(grepl("@MaximeBernier",text)) %>% 
  #filter(name == "Justin Trudeau") %>% 
  select(hashtags) %>% 
  #filter(is_retweet ==FALSE) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hastags in tweets mentioning Bernier",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )
  
#Top 10 hastags in tweets mentioning Scheer
canpoli.search.complete %>%
   filter(grepl("@AndrewScheer",text)) %>% 
  #filter(name == "Justin Trudeau") %>% 
  select(hashtags) %>% 
  #filter(is_retweet ==FALSE) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hastags in tweets mentioning Scheer",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )
  
#Top 10 hastags in tweets mentioning Singh
canpoli.search.complete %>%
  filter(grepl("@theJagmeetSingh",text)) %>% 
  select(hashtags) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hastags in tweets mentioning Singh",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )

#Top 10 hastags in tweets mentioning May
canpoli.search.complete %>%
  filter(grepl("@ElizabethMay",text)) %>% 
  select(hashtags) %>% 
  mutate(tolower(hashtags)) %>% 
  unnest_tokens(word, hashtags) %>%
  count(word, sort = TRUE) %>%
  filter(!is.na(word), n>1) %>%
  ungroup() %>% 
  arrange(desc(n)) %>% 
  top_n(10) %>% 
  ggplot(., aes(x = reorder(word, -n),y=n, fill=n)) +
  geom_bar(stat='identity')+
  theme(legend.title = element_blank())+
  coord_flip()+
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Top 10 hastags in tweets mentioning May",
    subtitle = "Tweet counts:  February 18 - March 29, 2019",
    caption = "\nSource: Twitter REST API via rtweet"
  )
```

##Correlations

```{r corr}
corr=tibble("party"=c("con","lib","ndp","grn","ppc"),"poll_pos" =1:5, "tweets_ot"=c(1,4,5,2,3), "tweets_all"=c(2,4,5,1,3), "followers"=c(3,1,4,2,5), "favourites"=c(2,1,4,5,3), "favourites_avg"=c(2,1,3,4,5))
chart.Correlation(corr[2:6], histogram = TRUE, method = "pearson")

```






